A meta-analysis of fungicide effects on mungbean powdery mildew

### 2021-05-19 04_Severity_MetaAnalysis.Rmd
add log of pecent severity meta-analysis, and plots of the residuals

# scale to 100% 
# create column giving the respective disease severity in the control 
PM_dat_D$yi_Cp <- unlist(apply(PM_dat_D, 1, function(x1){

@@

#logit transformed
PM_dat_D$yi_logit <- log(
   (PM_dat_D$yi_percent) / 
      (1 - (PM_dat_D$yi_percent))
   )
hist(PM_dat_D$yi_logit)
PM_dat_D$vi_logit <-
   (PM_dat_D$vi_percent) /
   (PM_dat_D$yi_percent * (1 - PM_dat_D$yi_percent)) ^ 2

@@

## Disease severity spray schedule meta-analysis  
First lets look at the results of the meta-analysis without transformation
```{r Metafor-Sev-analysis, cache=TRUE}
PMsev_mv <- rma.mv(
   yi = PM_final_severity,
   vi,
   mods = ~ spray_management,
   method = "ML",
   random = list( ~ spray_management | trial, ~ 1 | id),
   struct = "UN",
   #control = list(optimizer = "optim"),
   data = PM_dat_D
)
summary(PMsev_mv)
```

@@

summary(PMsev_mvSQ)
```
This meta-analysis is more in-line with estimates from the yield meta-analysis. `Early` sprays are not effective at reducing powdery mildew severity and two or more sprays had higher efficacy than single sprays. Recommended sprays were the most effective spray timing at reducing disease severity.

Finally lets assess the percent severity
```{r Metafor-Sev-analysis, cache=TRUE}
# remove controls for response ratio meta-analysis
PMsev_mvPer <- rma.mv(
   yi = yi_logit,
   V = yi_logit,
   mods = ~ spray_management,
   method = "ML",
   random = list( ~ spray_management | trial, ~ 1 | id),
   struct = "UN",
   control = list(optimizer = "optim"),
   data = PM_dat_D
)

@@

#summary(PMsev_mvPer)
# Intecep <- coef(PMsev_mvPer)[1]
# ((coef(PMsev_mvPer)[2:6]+ Intecep)/100)* median(PM_dat_D$PM_final_severity) - (Intecep/100*median(PM_dat_D$PM_final_severity))
```
Model does not converge we will proceed with the square transformed severity
```{r}
#logit transformed
PM_dat_D$yi_logit <- log(
   (PM_dat_D$yi_percent) / 
      (1 - (PM_dat_D$yi_percent))
   )

>>>>

# create log transformed values
PM_dat_D$yi_log <- log(PM_dat_D$yi_percent)
PM_dat_D$vi_log <- (PM_dat_D$vi_percent) / (PM_dat_D$replicates * PM_dat_D$yi_percent^2)

PM_dat_D[PM_dat_D$yi_log == -Inf,"yi_log"] <- 0
PM_dat_D[PM_dat_D$vi_log == -Inf,"vi_log"] <- 0.001

# create column giving the respective disease severity in the control 
PM_dat_D$yi_Cp <- unlist(apply(PM_dat_D, 1, function(x1){

@@

#logit transformed percentages
PM_dat_D$yi_logit <- log(
   (PM_dat_D$yi_percent) / 
      (1 - (PM_dat_D$yi_percent))
   )
hist(PM_dat_D$yi_logit)
PM_dat_D$vi_logit <-
   (PM_dat_D$vi_percent) /
   (PM_dat_D$yi_percent * (1 - PM_dat_D$yi_percent)) ^ 2

@@

## Disease severity spray schedule meta-analysis  
First lets look at the results of the meta-analysis without transformation
```{r Metafor-Sev-analysis, cache=TRUE}
PMsev_mv <- rma.mv(
   yi = PM_final_severity,
   vi,
   mods = ~ spray_management,
   method = "ML",
   random = list( ~ spray_management | trial, ~ 1 | id),
   struct = "UN",
   #control = list(optimizer = "optim"),
   data = PM_dat_D
)
summary(PMsev_mv)
plot(rstandard(PMsev_mv)$z, ylim=c(-3,3), pch=19)
```

@@

summary(PMsev_mvSQ)
plot(rstandard(PMsev_mvSQ)$z, ylim=c(-3,3), pch=19)
```
This meta-analysis is more in-line with estimates from the yield meta-analysis. `Early` sprays are not effective at reducing powdery mildew severity and two or more sprays had higher efficacy than single sprays. Recommended sprays were the most effective spray timing at reducing disease severity.

Finally lets assess the logit transformed percent severity
```{r Metafor-Sev-analysis_log, cache=TRUE}
# remove controls for response ratio meta-analysis
PMsev_mv_log <- rma.mv(
   yi = yi_log,
   V = yi_log,
   mods = ~ spray_management,
   method = "ML",
   random = list( ~ spray_management | trial, ~ 1 | id),
   struct = "US",
   control = list(optimizer = "optim"),
   data = PM_dat_D
)

summary(PMsev_mv_log)
plot(rstandard(PMsev_mv_log)$z, ylim=c(-3,3), pch=19)
data.frame(cbind((1-exp(PMsev_mv_log$b)),
                              (1-exp(PMsev_mv_log$ci.lb)),
                              (1-exp(PMsev_mv_log$ci.ub))))
```

Finally lets assess the logit transformed percent severity
```{r Metafor-Sev-analysis_logitP, cache=TRUE}
# remove controls for response ratio meta-analysis
PMsev_mvPer <- rma.mv(
   yi = yi_logit,
   V = yi_logit,
   mods = ~ spray_management,
   method = "ML",
   random = list( ~ spray_management | trial, ~ 1 | id),
   struct = "UN",
   control = list(optimizer = "optim"),
   data = PM_dat_D
)

@@

1- LaplacesDemon::invlogit(PMsev_mvPer$b)
mean(PM_dat_D[PM_dat_D$spray_management == "control","PM_final_severity"])
mean(PM_dat_D[PM_dat_D$spray_management == "Recommended_plus","PM_final_severity"])
```
This produces estimates which are too high and likely incorrect

```{r Metafor-Sev-analysis_logit, cache=TRUE}
#logit transformed 1- 9 values
PM_dat_D$yi_logit <- log(
   ((PM_dat_D$PM_final_severity-1)/8) / 
      (1 - ((PM_dat_D$PM_final_severity-1)/8))
   )

PM_dat_D$vi_logit <-
   (PM_dat_D$vi) /
   (((PM_dat_D$PM_final_severity-1)/8) * (1 -((PM_dat_D$PM_final_severity-1)/8))) ^ 2

# Correct -Inf to zeros
PM_dat_D[PM_dat_D$yi_logit == -Inf |
            PM_dat_D$yi_logit == Inf ,"yi_logit"] <- 0
PM_dat_D[is.na(PM_dat_D$vi_logit == Inf),"vi_logit"] <- 0

PM_dat_D[PM_dat_D$vi_logit <= 0,"vi_logit"] <- 0.001
# remove controls for response ratio meta-analysis
PMsev_mvPer <- rma.mv(
   yi = yi_logit,
   V = yi_logit,
   mods = ~ spray_management,
   method = "ML",
   random = list( ~ spray_management | trial, ~ 1 | id),
   struct = "UN",
   control = list(optimizer = "optim"),
   data = PM_dat_D
)

summary(PMsev_mvPer)
```

```{r}
#logit transformed
PM_dat_D$yi_logit <- log(
   (PM_dat_D$yi_percent) / 
      (1 - (PM_dat_D$yi_percent))
   )
